---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
suppressPackageStartupMessages(library(ezHawkSQL))
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ezHawkSQL

<!-- badges: start -->
<!-- badges: end -->

ezHawkSQL is intended to make querying from R in an integrated security environment easier. It contains a bunch of helper functions that help you manage referenced DSNs during an R session and help you save DSNs to your .Renviron. It also contains a set of functions that help you query SQL databases using the aforementioned DSNs. The Hawk in the name reflects the author's current favorite community college mascot. Go Hawks!

## Installation

You can install the development version of ezHawkSQL from [GitHub](https://github.com/) with:

``` {r}
# install.packages("devtools")
devtools::install_github("paulmeinz/ezHawkSQL")
```

## Setting your DSNs
Upon start-up and initial attachment of this package, you may receive a notification that you have no identified DSNs. That's a bummer! However, the ezHawkDSN package provides a number of helper functions to make this easy.

The first thing you'll want to do is use the set_dsn() function within your R session. This function takes two arguments - dsn and default. Here dsn should be a named character vector. Each item of this vector should be a data source name from you current windows environment. The names of these vector will act as aliases for your DSN (for convenience purposes) and are therefore heretofore referred to as *aliases* 

```{r}
dsns <- c('foo' = 'bar', 'spam' = 'beautiful_spam')
set_dsn(dsn = dsns, default = 'spam')
```

Then you can use get_dsn() to display the current dsns in your session.

```{r}
get_dsn()
```

And finally, the data source names you set during your session will not be saved for the next session. If you are hoping to have the convenience of entering your DSNs one time, you'll need to make appropriate edits to your .Renviron file. This is easy with ezHawkSQL! Just use the save_dsn() function (which takes no arguments) and follow the instructions.

```{r eval = FALSE}
save_dsn()
```

## Querying made easy
Now that you've worked out your DSN situation, you can use the SQL helper functions that I've written to make SQL EZ. The first helper function allows you to read queries - read_query(). This function can either read a single select query or it can read a multi-step query. The multi-step query must begin with the execution of several temp table steps (and/or other steps that do not return a data set) and end with a select statement that pulls the data. Each step should be divided by '---step---' so the function can easily distinguish between SQL steps. In the event of a multi-step query, the function will return a named vector. Each element will be a step in the query (in it's relative order of execution), and each element will be provided a name based on the order of execution (e.g., "Step 1", "Step 2", etc.) 

```{r eval = FALSE}
my_sql_query_directory <- 'spam_sql.txt'
query <- read_query(my_sql_query_directory)
```

Next you can use sql_query() to gather data from your favorite data source. This requires two arguments - a query and an alias. If you do not provide an alias, sql_query will use your default alias. This means you only need to provide a query! How neat.

```{r eval = FALSE}
sql_query("select spam from beautiful_spam where spam = 'yum'", 'spam')

# OR
my_sql_query_directory <- 'spam_sql.txt'
query <- read_query(my_sql_query_directory)
sql_query(query)
```

sql_query() query will however not work super well for a multi-step query - e.g. a query that has several preparatory steps prior to pulling the final data set. For that, you can use multistep_sql_query(). This function takes as it's first argument a named list of queries in the order of execution (e.g. first element gets executed first, etc.). The second argument is your DSN alias.

```{r eval = FALSE}
step1 <- "select spam into #spam_table from beautiful_spam"
step2 <- "select * from #spam_table"
queries <- c('Step 1' = step1, 'Step 2' = step2)
multistep_sql_query(queries)
```

